---
title: "TimeBasedData"
author: "Anusha Gadgil"
date: "2023-11-07"
output: html_document
---
##Importing the Dataset and Necessary Libraries
```{r }
library(tidyverse)
library(dplyr)
library(reshape2)
library(gridExtra)
library(ggplot2)
```

##Importing the Dataset

```{r}
WineDataset=read.csv("C:/Users/Anusha Gadgil/Desktop/wines_SPA (1).csv")
WineDataset
```

```{r}
library(tsibble)
library(dplyr)

# Assuming WineDataset is your data frame

# Convert year to Date class
#WineDataset$year <- as.Date(paste0(WineDataset$year))

# Create tsibble
#my_tsibble <- as_tsibble(WineDataset, key = "year")

# Print the tsibble
#print(my_tsibble)
my_tsibble <- my_tsibble %>%
  mutate(Time = as.Date(Time))

```

#Selecting and converting year column to timebased format
```{r}
WineDataset$year<-as.Date(as.character(WineDataset$year), format="%Y")
WineDataset
```
#checking total types of wine

```{r}
total_types <- WineDataset %>%
  summarise(total_types = n_distinct(type))
total_types
```

#grouping together wine of same type and year and analysing price change
```{r}
library(dplyr)
#converting year from int to date
WineDataset$year <- as.Date(WineDataset$year)

# Grouping together wines of the same type and calculating the average price
grouped_data <- WineDataset %>%
  group_by(type,year) %>%
  summarise(average_price = mean(price, na.rm = TRUE))

# Print the result
print(grouped_data)

```


#Creating a Tsibble
```{r}
df_unique<- na.omit(grouped_data)
#duplicates <- duplicated(grouped_data)
#df_unique <- grouped_data[!duplicates, ]
#df_unique
df_unique
```

```{r}
redundant <- duplicated(df_unique)

print(redundant)

```

```{r}
df_unique$year <- as.Date(df_unique$year, format = "%d-%m-%Y")
df_unique
```



```{r}
my_tsibble_no_na <- na.omit(my_tsibble)
# Create tsibble
my_tsibble <- tsibble(
  Time = df_unique$year,
  Type = df_unique$type,
  Average_Price = df_unique$average_price,
  key = c(Type)
)
# Print the tsibble
print(my_tsibble)

```


#analysing price of Rioja White over time
```{r}
filtered_tibble<-filter(my_tsibble,Type=="Rioja White")

ggplot(filtered_tibble, aes(x = Time, y = Average_Price)) +
  geom_line() +
  labs(title = "Total Sales over Time", x = "Time", y = "Price of Wine") +
  theme_minimal()
```

#analysing price of Albarino over time
```{r}
filtered_tibble<-filter(my_tsibble,Type=="Albarino")

ggplot(filtered_tibble, aes(x = Time, y = Average_Price)) +
  geom_line() +
  labs(title = "Total Sales over Time", x = "Time", y = "Price of Wine") +
  theme_minimal()
```
#Analysing price of Montsant Red over time
```{r}
filtered_tibble<-filter(my_tsibble,Type=="Montsant Red")

ggplot(filtered_tibble, aes(x = Time, y = Average_Price)) +
  geom_line() +
  labs(title = "Total Sales over Time", x = "Time", y = "Price of Wine") +
  theme_minimal()

```



#All wine prices showed a downward trend over time

```{r}
trend_results <- my_tsibble %>%
  group_by(Type) %>%
  do(
    model = lm(Average_Price ~ Time, data = .)
  )

# Extract coefficients (slope) from the linear models
trend_results <- trend_results %>%
  summarise(
    Type = unique(Type),
    Slope = coef(model)[2]  # Extract the slope (coefficient) for Time
  )

# Print the trend results
print(trend_results)

```
```{r}

summary(my_tsibble)


```


#every year the price of the wine is constantly decreasing. The trends are not that strong as the decrease in price is not that much.

```{r}
library(tsibble)
library(dplyr)
library(fable)
df$Time <- as.Date(df$Time)
my_tsibble <- tsibble(Time = df$Time, Type = df$Type, Average_Price = df$Average_Price)
                                       
                                       
# Apply stl decomposition
smoothed_data <- my_tsibble %>%
  model(stl(Average_Price ~ season("year")))


```

